---
title: "Project_1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project_1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(LAGG4793)
welding.csv = LAGG4793::project_data
welding.csv$X = NULL
```

# Introduction
The following project is investigating the welding data set and how robotic welders stay within a range of designated parameters while welding drive shafts of automobiles. In this project I will be looking at the normality of the data set. Then I will use control charts, control ellipses, and univariate $\bar x$ charts to determine if the automated welding is sufficient. 

## Importance
As mentioned I am investigating if the robotic welders stay within the designated parameters. This is an important problem, because if the welders are not staying withing their designated parameters they will not produce a quality product. We can use $T^2$ charts, control ellipses, and $\bar x$ charts to determine points that are far outside the parameters and then the engineers can determine what went wrong.

## Data
The variables that are being investigated are as follow

voltage- measured in volts

current- measured in amps

feed speed- measured in inches per minute

gas flow- measured in cubic feet per minute

The first six entries of the data is below
```{r}
head(LAGG4793::project_data)
```
Now that we have a good understanding of the data we can begin normality checks

# Checking Normailty
We will first check the normality of the data using three different methods.

1. Univariate check

2. Multivariate check

3. Chi-sq QQ plot

If the data is not normal then we will do a log transformation and see if that passes the normality checks

## Univariate check
To do a univariate check of the data we will plot each variable versus its quantile and run a shapiro-wilk test using the univariate_check function
```{r}
p_values = LAGG4793::univariate_check(welding.csv)
p_values
```

From the plots and the output of the function we see that the voltage, current, and feed speed variables all pass the shapiro-wilk test, but the gas flow variable does not pass the test. As a result we will take the log of gas flow and re-run the shapiro-wilk test.

### Log univariate check
```{r}
welding_log = data.frame("logGas" = log(welding.csv$Gasf))
univariate_check(welding_log)
```

From this transformation we see that a log transformation does not really do much to the data, and this is not surprising since many of the datapoints share the same values.


## Bivariate Check
I will now do a bivariate check. This is done by making a confidence ellipse between two variables at alpha = .5 and alpha = 0.05
```{r}
LAGG4793::bivariate_check(welding.csv, 0.5)
```

From the graphs and the output of the functions we can see that roughly half of the sample data points lie within the ellipse. And although the voltage vs feedspd ellipse only has a proportion of 0.425 inside the ellipse this may just be due to the sample being too small.

```{r}
LAGG4793::bivariate_check(welding.csv, 0.05)
```
From the graphs and the output of the functions we can see that roughly 95% of the sample data points lie within the ellipse as we would expect.


### Bivariate Check Log Transformation
Now I will do a log transformation on gas flow and run the same normality checks
```{r}
welding_log = welding.csv
welding_log$Gasf = log(welding.csv$Gasf)
```

Below is the alpha = 0.5 and alpha = 0.05
```{r}
LAGG4793::bivariate_check(welding_log, 0.5)
```

```{r}
LAGG4793::bivariate_check(welding_log, 0.05)
```
Unsurprisingly we get the same amount of points inside of the graph as outside of it

## Chi-sq QQ Plot
Now I will check the normality using the chi-squared QQ plot
```{r}
LAGG4793::chisq_plot(welding.csv)
```
From the above graph and the p-value of the shapiro wilk test we can see that the data fails the test of normality at an alpha of 0.05. Thus we are fairly confident that this data is not normal. To try and fix the data I will take the log transformation of the gas flow


### Chi-sq QQ Plot Log
```{r}
LAGG4793::chisq_plot(welding_log)
```
This data has a little higher p-value, but it still fails the test of normality at an alpha of 0.05.

## Normality Conclusion
As a conclusion we can say that for the voltage, current, and feed speed are normal independent of the other values. However, taking into account all of the variables together using the Chi-squared QQ plot we see that the data is not normal. Additionally, transforming the non-normal gas flow variable using a log transformation is not particularly effective in making the data normal. Nonetheless, we will still use the data where the log of gas flow is taken since the p value was a little higher in the Chi-sq QQ plot and because this is how it is done in the book, whose example I am following for this project.

# T-squared chart
Now I am going to make the $T^2$ chart for the data

```{r}
LAGG4793::Tsq_chart(welding_log)
```
From this plot we can see that observation 31 seems to be an outiler. Since it is out of control. 


## Investigating the Unusual Point
We will not investigate which variable seems to be causing this problem. To do this I will check the z scores of each variable
```{r}
m = colMeans(welding_log)
voltage_sd = sd(welding_log[,1])
current_sd = sd(welding_log[,2])
feed_sd = sd(welding_log[,3])
gas_sd = sd(welding_log[,4])


(welding_log[31,1] - m[1])/voltage_sd
(welding_log[31,2] - m[2])/current_sd
(welding_log[31,3] - m[3])/feed_sd
(welding_log[31,4] - m[4])/gas_sd



```
From the z scores we see that the gas feed variable is highly unusual thus we would recommend the engineers to look at this variable to see what went wrong.



# 99% Quality Control Ellipse
I will now make a log_gas_flow
```{r}
y = welding_log$Gasf
x = welding_log$Voltage

LAGG4793::quality_ellipse(0.01, y, x, "ln(gas_flow)", "Voltage")
```

Above is the 99% quality control ellipse. From this we see that all but one of the points lies within this ellipse meaning this point is unusual. 
And from the T-squared chart we know this point is point 31.
```{r}
welding_log[31,]
```
This is reconfirmed by the above output since this point has a log gas flow of greater than 4 which is the only point where this is true.





